---
title: "R refresher"
output: html_notebook
---
 
R is a language and environment for statistical computing and graphics....R provides a wide variety of statistical (linear and nonlinear modeling, classical statistical tests, time-series analysis, classification, clustering, ...) and graphical techniques, and is highly extensible.

```{r}
text <- "hello world"
print(text) # print is an R function. Here it prints text
# the hashtag comments out whatever is behin comes after it

```
variable types

Character: "learning"  
numeric: 3.5    
integer: 3  
logical: TRUE/FALSE

```{r}
#vectors can be defined in multiple ways
#using the vector() function
v <- vector("numeric", length = 5)

#using the c() function
w <- c(2,4) 
w.1 <- c(TRUE, FALSE)
w.2 <- c("a","b","c")
w.3 <- 3:7

w
w.1
w.2
w.3

```

List  
Special type of vector that can contain different elements
```{r}
l <- list("a", 4, TRUE)
l
```

Matrix
```{r}
#how to define an empty matrix of n rows and n columns
m <- matrix(nrow = 2, ncol = 4)

#creating matrix by binding columns
x <- 1:3
y <- 5:7
xy <- cbind(x,y)
yx <- rbind(x,y)

xy
yx
```

Data frames  
used to stores data with multiple rows and columns.   
they can stores different sort of data/elements (Numeric, Logical etc...).   
They are created by reading data of csv,excel or table format.   
They are also created by binding columns to each other or transforming matrices into dataframes using the dataframe() function.
```{r}
x  <- data.frame(return  =  c(2,-3,5,6,-5) , price  =  c("up","down", "up", "up", "down"))
x
```

We want to add a new column to X called "candletype"

```{r}
x$candletype <- rep(0,5)
```

Now we want to assign negative numbers in Price with a bear candle and positive numbers witha bull candle

```{r}
x$candletype <- ifelse(x$price < 0, "bear", "bull")
print(x)
```

Subsetting:  Say we want to filter the data based on some condition or we want to only select a few rows.
```{r}
#Select only the 1st 3 rows
x_3row <- x[1:3,]

#Select only the 1st 3 rows
x_2cols <- x[,1:2]

print(x_3row)
print(x_2cols)
```

